---
title: "Metrics for assessing community assembly processes"
output: html_notebook
bibliography: ../ref.bib  
---

```{r}
source('../env.R')
```

```{r}
community_data = read_csv(filename(COMMUNITY_OUTPUT_DIR, 'community_assembly_metrics_using_relative_abundance.csv'))
head(community_data)
colnames(community_data)
```

Join on realms
```{r}
city_to_realm = read_csv(filename(CITY_DATA_OUTPUT_DIR, 'realms.csv'))
community_data_with_realm = left_join(community_data, city_to_realm)
```

Cities as points
```{r}
city_points = st_centroid(read_sf(filename(CITY_DATA_OUTPUT_DIR, 'city_selection.shp'))) %>% left_join(community_data_with_realm)
city_points_coords = st_coordinates(city_points)
city_points$latitude = city_points_coords[,1]
city_points$longitude = city_points_coords[,2]
```
  
```{r}
world_map = read_country_boundaries()
```

Load community data, and create long format version
```{r}
communities = read_csv(filename(COMMUNITY_OUTPUT_DIR, 'communities_for_analysis.csv'))
communities
```
```{r}
community_summary = communities %>% group_by(city_id) %>% summarise(regional_pool_size = n(), urban_pool_size = sum(relative_abundance_proxy > 0))
community_summary
```

Load trait data
```{r}
traits = read_csv(filename(TAXONOMY_OUTPUT_DIR, 'traits_jetz.csv'))
head(traits)
```

Load realm geo
```{r}
resolve = read_resolve()
head(resolve)
```

# Summary metrics by Realm
```{r}
test_required_values = function(name, df) {
  cat(paste(
    test_value_wilcox(paste(name, 'MNTD'), df$mntd_normalised),
    test_value_wilcox(paste(name, 'Beak Gape FDiv'), df$gape_width_fdiv_normalised),
    test_value_wilcox(paste(name, 'HWI FDiv'), df$handwing_index_fdiv_normalised),
    test_value_wilcox(paste(name, 'Mass FDiv'), df$mass_fdiv_normalised),
    nrow(df),
    sep = "\n"))
}
```

```{r}
test_required_values('Global', community_data_with_realm)
```

```{r}
unique(community_data_with_realm$core_realm)
```

```{r}
test_required_values('Nearctic', community_data_with_realm[community_data_with_realm$core_realm == 'Nearctic',])
```

```{r}
test_required_values('Neotropic', community_data_with_realm[community_data_with_realm$core_realm == 'Neotropic',])
```

```{r}
test_required_values('Palearctic', community_data_with_realm[community_data_with_realm$core_realm == 'Palearctic',])
```

```{r}
test_required_values('Afrotropic', community_data_with_realm[community_data_with_realm$core_realm == 'Afrotropic',])
```

```{r}
test_required_values('Indomalayan', community_data_with_realm[community_data_with_realm$core_realm == 'Indomalayan',])
```

```{r}
test_required_values('Australasia', community_data_with_realm[community_data_with_realm$core_realm == 'Australasia',])
```

# Summary metrics by introduced species
```{r}
communities = read_csv(filename(COMMUNITY_OUTPUT_DIR, 'communities_for_analysis.csv'))
city_introduced_species = communities %>% group_by(city_id) %>% summarise(number_of_species = n()) %>% left_join(
  communities %>% group_by(city_id) %>% filter(origin == 'Introduced') %>% summarise(number_of_introduced_species = n())
) %>% replace_na(list(number_of_introduced_species = 0))

community_data_with_introductions = left_join(community_data, city_introduced_species)
community_data_with_introductions$has_introduced_species = community_data_with_introductions$number_of_introduced_species > 0
community_data_with_introductions
```
```{r}
community_data_with_introductions[,c('mntd_normalised', 'has_introduced_species')]
```

```{r}
community_data_with_introductions %>% group_by(has_introduced_species) %>% summarise(
  mean_mntd_normalised = mean(mntd_normalised, na.rm = T),
  sd_mntd_normalised = sd(mntd_normalised, na.rm = T),
  
  mean_mass_fdiv_normalised = mean(mass_fdiv_normalised, na.rm = T),
  sd_mass_fdiv_normalised = sd(mass_fdiv_normalised, na.rm = T),
  
  mean_gape_width_fdiv_normalised = mean(gape_width_fdiv_normalised, na.rm = T),
  sd_gape_width_fdiv_normalised = sd(gape_width_fdiv_normalised, na.rm = T),
  
  mean_handwing_index_fdiv_normalised = mean(handwing_index_fdiv_normalised, na.rm = T),
  sd_handwing_index_fdiv_normalised = sd(handwing_index_fdiv_normalised, na.rm = T)
)
```

## MNTD
```{r}
ggplot(community_data_with_introductions, aes(x = has_introduced_species, y = mntd_normalised)) + geom_boxplot()
```

```{r}
wilcox.test(mntd_normalised ~ has_introduced_species, community_data_with_introductions)
```

There is a significant difference between the response of cities with introduced species (0.53±0.27) and those without (0.47±0.19) (p-value = 0.02).


## Mass FDiv
```{r}
ggplot(community_data_with_introductions, aes(x = has_introduced_species, y = mass_fdiv_normalised)) + geom_boxplot()
```

```{r}
wilcox.test(mass_fdiv_normalised ~ has_introduced_species, community_data_with_introductions)
```
There is a significant difference between the response of cities with introduced species (0.57±0.27) and those without (0.73±0.24) (p < 0.0001)


## Beak Gape FDiv
```{r}
ggplot(community_data_with_introductions, aes(x = has_introduced_species, y = gape_width_fdiv_normalised)) + geom_boxplot()
```

```{r}
wilcox.test(gape_width_fdiv_normalised ~ has_introduced_species, community_data_with_introductions)
```
There is NOT a significant difference between the response of cities with introduced species (0.61±0.30) and those without (0.56±0.27)


## HWI FDiv
```{r}
ggplot(community_data_with_introductions, aes(x = has_introduced_species, y = handwing_index_fdiv_normalised)) + geom_boxplot()
```

```{r}
wilcox.test(handwing_index_fdiv_normalised ~ has_introduced_species, community_data_with_introductions)
```
There is a significant difference between the response of cities with introduced species (0.49±0.30) and those without (0.79±0.21) (p < 0.0001)


# Examine individual metrics

## Analysis data frame
```{r}
analysis_data = community_data_with_realm[,c('city_id', 'mntd_normalised', 'mass_fdiv_normalised', 'gape_width_fdiv_normalised', 'handwing_index_fdiv_normalised', 'core_realm')] %>% 
  left_join(city_points[,c('city_id', 'latitude', 'longitude')]) %>%
  left_join(community_data_with_introductions[,c('city_id', 'has_introduced_species')])
```

## Helper plot functions
```{r}
geom_normalised_histogram = function(name, gg, legend.position = "right") {
  gg + 
    geom_histogram(aes(fill = core_realm), binwidth = 0.1, position = "dodge") +
    geom_vline(aes(xintercept = 0.5), color = "#000000", size = 0.4) +
    geom_vline(aes(xintercept = 0), color = "#000000", size = 0.2, linetype = "dashed") +
    geom_vline(aes(xintercept = 1), color = "#000000", size = 0.2, linetype = "dashed") + 
    ylab("Number of cities") + xlab("Normalised Response") + ylim(c(0, 70)) +
    labs(title = name, fill = 'Realm') +
    theme_bw() +
    theme(legend.position=legend.position)
}
```

```{r}
geom_map = function(map_sf, title) {
  norm_mntd_analysis_geo = ggplot() + 
    geom_sf(data = world_map, aes(geometry = geometry)) +
    map_sf +
    normalised_colours_scale +
    labs(title = title, colour = 'Normalised\nResponse')
}
```

## MNTD
```{r}
norm_mntd_analysis_plot = geom_normalised_histogram(
  'MNTD', 
  ggplot(analysis_data, aes(mntd_normalised))
)
norm_mntd_analysis_plot
```

```{r}
norm_mntd_analysis_geo = geom_map(geom_sf(data = analysis_data, aes(color = mntd_normalised, geometry = geometry)), 'MNTD')
norm_mntd_analysis_geo
ggsave(filename(FIGURES_OUTPUT_DIR, 'normalised_mntd_using_abundance.jpg'), width = 2500, height=1100, units = 'px')
```

```{r}
summary(lm(data = analysis_data, mntd_normalised ~ latitude))
```

```{r}
summary(lm(data = analysis_data, mntd_normalised ~ abs(latitude)))
```

## Gape width - FDiv
```{r}
norm_gape_fdiv_analysis_plot = geom_normalised_histogram(
  'Gape FDiv', 
  ggplot(analysis_data, aes(gape_width_fdiv_normalised))
)
norm_gape_fdiv_analysis_plot
```

```{r}
norm_gape_fdiv_analysis_geo = geom_map(geom_sf(data = analysis_data, aes(color = gape_width_fdiv_normalised, geometry = geometry)), 'Gape Width FDiv')
norm_gape_fdiv_analysis_geo
ggsave(filename(FIGURES_OUTPUT_DIR, 'normalised_gape_width_using_abundance.jpg'), width = 2500, height=1100, units = 'px')
```

```{r}
summary(lm(data = analysis_data, gape_width_fdiv_normalised ~ latitude))
```

```{r}
summary(lm(data = analysis_data, gape_width_fdiv_normalised ~ abs(latitude)))
```

## HWI - FDiv
```{r}
norm_hwi_fdiv_loco_analysis_plot = geom_normalised_histogram(
  'HWI FDiv', 
  ggplot(analysis_data, aes(handwing_index_fdiv_normalised))
)
norm_hwi_fdiv_loco_analysis_plot
```

```{r}
norm_hwi_fdiv_analysis_geo = geom_map(geom_sf(data = analysis_data, aes(color = handwing_index_fdiv_normalised, geometry = geometry)), 'HWI FDiv')
norm_hwi_fdiv_analysis_geo
ggsave(filename(FIGURES_OUTPUT_DIR, 'normalised_hwi_using_abundance.jpg'), width = 2500, height=1100, units = 'px')
```

```{r}
summary(lm(data = analysis_data, handwing_index_fdiv_normalised ~ latitude))
```

```{r}
summary(lm(data = analysis_data, handwing_index_fdiv_normalised ~ abs(latitude)))
```

# Compare metrics against each other
```{r}
ggplot(analysis_data, aes(x = gape_width_fdiv_normalised, y = mntd_normalised, colour = core_realm)) + 
  geom_point() +
  ylab("MNTD") + 
  xlab("Gape Width FDiv") +
  theme_bw() + labs(color = "Realm")
```

```{r}
ggplot(analysis_data, aes(x = handwing_index_fdiv_normalised, y = mntd_normalised, colour = core_realm)) + 
  geom_point() +
  ylab("MNTD") + 
  xlab("HWI FDiv") +
  theme_bw() + labs(color = "Realm")
```

```{r}
ggplot(analysis_data, aes(x = handwing_index_fdiv_normalised, y = gape_width_fdiv_normalised, colour = core_realm)) + 
  geom_point() +
  ylab("Gape Width FDiv") + 
  xlab("HWI FDiv") +
  theme_bw() + labs(color = "Realm")
```

```{r}
mntd_fdiv_analysis = analysis_data %>% 
  dplyr::select(city_id,  mntd_normalised, handwing_index_fdiv_normalised, gape_width_fdiv_normalised) %>%
  left_join(community_summary) %>%
  mutate(urban_pool_perc = urban_pool_size * 100 / regional_pool_size)
mntd_fdiv_analysis
```

```{r}
ggpairs(mntd_fdiv_analysis %>% dplyr::select(mntd_normalised, handwing_index_fdiv_normalised, gape_width_fdiv_normalised, regional_pool_size, urban_pool_size, urban_pool_perc))
ggsave(filename(FIGURES_OUTPUT_DIR, 'appendix_normalised_correlation.jpg'))
```


