---
title: "Create regional pools using BirdLife"
output: html_notebook
bibliography: ../ref.bib 
---

```{r}
source('../env.R')
```

# Generate Worldwide Distributions for all Columbidae

## Read in the Birdlife Distribution 
Birdlife dataset @Birdlife2020.

```{r}
birdlife = st_read(DL_BIRDLIFE_DISTRIBUTIONS)
head(birdlife)
```

## Read in the Birdlife Taxonomy
Load the birdlife taxonomy to find Columbidae species.

```{r}
st_crs(birdlife)
```

```{r}
birdlifeTaxonomyWb = read_excel(DL_BIRDLIFE_TAXONOMY)
birdlifeTaxonomyWb
```

```{r}
names(birdlifeTaxonomyWb) <- birdlifeTaxonomyWb[2,]
birdlifeTaxonomy <- birdlifeTaxonomyWb[-2,]
birdlifeTaxonomy
```
Fetch all columbidae species that are not extinct and are not an invalid birdlife species.
```{r}
birdlifeColumbidae = birdlifeTaxonomy[!is.na(birdlifeTaxonomy$Order) & birdlifeTaxonomy$Order == 'COLUMBIFORMES', c('SISRecID', 'Scientific name', 'Common name', 'Synonyms', '2023 IUCN Red List category')]
names(birdlifeColumbidae) = c('birdlife_id', 'species_name', 'common_name', 'synonyms', 'red_list_2023')
birdlifeColumbidae = birdlifeColumbidae[!(birdlifeColumbidae$red_list_2023 %in% c('EX', 'NR')),]
birdlifeColumbidae
```

Check to see if we have any missing distributions
```{r}
birdlifeColumbidae[!(birdlifeColumbidae$species_name %in% birdlife$SCI_NAME),]
```

## Store Birdlife taxonomy
```{r}
write_csv(birdlifeColumbidae, filename(BIRDLIFE_WORKING_OUTPUT_DIR, 'taxonomy.csv'))
birdlifeColumbidae = read_csv(filename(BIRDLIFE_WORKING_OUTPUT_DIR, 'taxonomy.csv'))
head(birdlifeColumbidae)
```

## Filter Birdlife Distributions to only Columbidae
Do we have any distributions not in our taxonomy, are are not extinct?
```{r}
birdlife[!(birdlife$SCI_NAME %in% birdlifeColumbidae$species_name) & birdlife$PRESENCE != 5,]
```

```{r}
COLUMBIDAE_WORLD_DISTRIBUTIONS = filename(BIRDLIFE_WORKING_OUTPUT_DIR, "columbidae_world.shp")
write_sf(birdlife[birdlife$PRESENCE != 5,], COLUMBIDAE_WORLD_DISTRIBUTIONS)
columbidaeDistribution = read_sf(COLUMBIDAE_WORLD_DISTRIBUTIONS)
```


Free up memory by removing birdlife data from workspace
```{r}
rm(birdlife)
rm(birdlifeTaxonomy)
rm(birdlifeTaxonomy_tmp)
rm(birdlifeTaxonomyWb)
rm(birdlifeTaxonomyWb_Sheets)
```

```{r}
st_crs(columbidaeDistribution)
```

# Create intersection of Columbidae Distributions with Urban Spaces
Read in our initial city vectors, we will join these to our checklists
```{r}
initial_city_selection = read_sf(filename(mkdir(GEO_WORKING_OUTPUT_DIR, 'cities'), 'initial_selection.shp'))
```

```{r}
st_crs(initial_city_selection)
```

```{r}
urban_columbidae_distributions = st_join(initial_city_selection, columbidaeDistribution) 

urban_columbidae_distributions
```

## Check birdlife distribution data

Species with unknown presence?
```{r}
urban_columbidae_distributions[urban_columbidae_distributions$PRESENCE > 5, c('SCI_NAME', 'city_name', 'city_id')]
```



Possibly Extinct species
```{r}
urban_columbidae_distributions[urban_columbidae_distributions$PRESENCE == 4, c('SCI_NAME', 'city_name', 'city_id')]
```

## Map seasonal and origin to string values
```{r}
urban_columbidae_distributions$seasonal_desc = "Unknown Code"
urban_columbidae_distributions$seasonal_desc[urban_columbidae_distributions$SEASONAL == "1"] = "Resident"
urban_columbidae_distributions$seasonal_desc[urban_columbidae_distributions$SEASONAL == "2"] = "Breeding Season" 
urban_columbidae_distributions$seasonal_desc[urban_columbidae_distributions$SEASONAL == "3"] = "Non-breeding Season" 
urban_columbidae_distributions$seasonal_desc[urban_columbidae_distributions$SEASONAL == "4"] = "Passage" 
urban_columbidae_distributions$seasonal_desc[urban_columbidae_distributions$SEASONAL == "5"] = "Seasonal occurence uncertain"
```

```{r}
urban_columbidae_distributions$origin_desc = "Unknown Code"
urban_columbidae_distributions$origin_desc[urban_columbidae_distributions$ORIGIN == "1"] = "Native"
urban_columbidae_distributions$origin_desc[urban_columbidae_distributions$ORIGIN == "2"] = "Reintroduced" 
urban_columbidae_distributions$origin_desc[urban_columbidae_distributions$ORIGIN == "3"] = "Introduced" 
urban_columbidae_distributions$origin_desc[urban_columbidae_distributions$ORIGIN == "4"] = "Vagrant" 
urban_columbidae_distributions$origin_desc[urban_columbidae_distributions$ORIGIN == "5"] = "Origin Uncertain"
urban_columbidae_distributions$origin_desc[urban_columbidae_distributions$ORIGIN == "6"] = "Assisted Colonisation"
```

```{r}
urban_columbidae_distributions$presence_desc = "Unknown Code"
urban_columbidae_distributions$presence_desc[urban_columbidae_distributions$PRESENCE == "1"] = "Extant"
urban_columbidae_distributions$presence_desc[urban_columbidae_distributions$PRESENCE == "2"] = "Probably Extant" 
urban_columbidae_distributions$presence_desc[urban_columbidae_distributions$PRESENCE == "3"] = "Possibly Extant" 
urban_columbidae_distributions$presence_desc[urban_columbidae_distributions$PRESENCE == "4"] = "Possibly Extinct" 
urban_columbidae_distributions$presence_desc[urban_columbidae_distributions$PRESENCE == "5"] = "Extinct"
urban_columbidae_distributions$presence_desc[urban_columbidae_distributions$PRESENCE == "6"] = "Presence Uncertain"
```

```{r}
cache_urban_columbidae_distributions = filename(BIRDLIFE_WORKING_OUTPUT_DIR, 'urban_distributions_birdlife_only.csv')
write_csv(urban_columbidae_distributions, cache_urban_columbidae_distributions)
urban_columbidae_distributions = read_csv(cache_urban_columbidae_distributions)
```

## Missing North American distribution of Streptopelia decaocto and Columba Livia
Birdlife is missing the recent invasion of Streptopelia decaocto into North America. 
We supplement the data for Streptopelia decaocto with the range from eBird's Status and Trends dataset.

```{r}
world_map = read_country_boundaries()
```
### Collared Dove
Existing range from Birdlife
```{r}
coldov_vect = st_simplify(columbidaeDistribution[columbidaeDistribution$SCI_NAME == 'Streptopelia decaocto', c('PRESENCE')], dTolerance = 0.02)

ggplot() +  
  geom_sf(data = world_map, aes(geometry = geometry)) +
  geom_sf(data = coldov_vect, aes(geometry = geometry), fill = "red")
```

Range from eBird
```{r}
eucdov_range_vect = st_read(DL_EBIRD_EUCDOV_RANGE)

ggplot() +  
  geom_sf(data = world_map, aes(geometry = geometry)) +
  geom_sf(data = eucdov_range_vect["scientific_name"], aes(geometry = geom), fill = "red")
```

Extract cities from eBird range
```{r}
existing_eucdov_cities = urban_columbidae_distributions$city_id[urban_columbidae_distributions$SCI_NAME == 'Streptopelia decaocto']

eucdov_distribution = st_join(initial_city_selection, eucdov_range_vect) 

eucdov_distribution[!is.na(eucdov_distribution$species_code) & !(eucdov_distribution$city_id %in% existing_eucdov_cities),]
```


```{r}
missing_eucdov_cities = data.frame(
    species_name = 'Streptopelia decaocto', 
    city_id = eucdov_distribution$city_id[!is.na(eucdov_distribution$species_code) & !(eucdov_distribution$city_id %in% existing_eucdov_cities)],
    city_name= eucdov_distribution$city_name[!is.na(eucdov_distribution$species_code) & !(eucdov_distribution$city_id %in% existing_eucdov_cities)],
    origin = 'Introduced',
    seasonal = 'Resident',
    presence = 'Extant'
)
missing_eucdov_cities
```

### Rock dove
Existing range from Birdlife
```{r}
rocdov_vect = st_simplify(columbidaeDistribution[columbidaeDistribution$SCI_NAME == 'Columba livia', c('PRESENCE')], dTolerance = 0.02)

ggplot() +  
  geom_sf(data = world_map, aes(geometry = geometry)) +
  geom_sf(data = rocdov_vect, aes(geometry = geometry), fill = "red")
```

Range from eBird
```{r}
rocpig_range_vect = st_read(DL_EBIRD_ROCPIG_RANGE)
ggplot() +  
  geom_sf(data = world_map, aes(geometry = geometry)) +
  geom_sf(data = rocpig_range_vect["scientific_name"], aes(geometry = geom), fill = "red")
```

```{r}
existing_rocpig_cities = urban_columbidae_distributions$city_id[urban_columbidae_distributions$SCI_NAME == 'Columba livia']

rocpig_distribution = st_join(initial_city_selection, rocpig_range_vect) 

rocpig_distribution[!is.na(rocpig_distribution$species_code) & !(rocpig_distribution$city_id %in% existing_rocpig_cities),]
```

```{r}
missing_rocpig_cities = data.frame(
    species_name = 'Columba livia', 
    city_id = rocpig_distribution$city_id[!is.na(rocpig_distribution$species_code) & !(rocpig_distribution$city_id %in% existing_rocpig_cities)],
    city_name= rocpig_distribution$city_name[!is.na(rocpig_distribution$species_code) & !(rocpig_distribution$city_id %in% existing_rocpig_cities)],
    origin = 'Introduced',
    seasonal = 'Resident',
    presence = 'Extant'
)
missing_rocpig_cities
```

## Add Spotted dove's expanded range into China
Existing range from Birdlife
```{r}
spodov_vect = st_simplify(columbidaeDistribution[columbidaeDistribution$SCI_NAME %in% c('Spilopelia chinensis','Spilopelia suratensis','Patagioenas albipennis'), c('SCI_NAME')], dTolerance = 0.02)

ggplot() +  
  geom_sf(data = world_map, aes(geometry = geometry)) +
  geom_sf(data = spodov_vect, aes(geometry = geometry, fill = SCI_NAME))
```

Range from eBird
```{r}
spodov_range_vect = st_read(DL_EBIRD_SPODOV_RANGE)
ggplot() +  
  geom_sf(data = world_map, aes(geometry = geometry)) +
  geom_sf(data = spodov_range_vect["scientific_name"], aes(geometry = geom), fill = "red")
```

```{r}
existing_spodov_cities = urban_columbidae_distributions$city_id[urban_columbidae_distributions$SCI_NAME %in% c('Spilopelia chinensis','Spilopelia suratensis','Patagioenas albipennis')]

spodov_distribution = st_join(initial_city_selection, spodov_range_vect) 

spodov_distribution[!is.na(spodov_distribution$species_code) & !(spodov_distribution$city_id %in% existing_spodov_cities),]
```

```{r}
missing_spodov_cities = data.frame(
    species_name = 'Spilopelia chinensis', 
    city_id = spodov_distribution$city_id[!is.na(spodov_distribution$species_code) & !(spodov_distribution$city_id %in% existing_spodov_cities)],
    city_name= spodov_distribution$city_name[!is.na(spodov_distribution$species_code) & !(spodov_distribution$city_id %in% existing_spodov_cities)],
    origin = 'Introduced',
    seasonal = 'Resident',
    presence = 'Extant'
)
missing_spodov_cities
```

## Store regional pool result
```{r}
urban_distribution = rbind(
  data.frame(
    species_name = urban_columbidae_distributions$SCI_NAME, 
    city_id = urban_columbidae_distributions$city_id,
    city_name= urban_columbidae_distributions$city_name,
    origin = urban_columbidae_distributions$origin_desc,
    seasonal = urban_columbidae_distributions$seasonal_desc,
    presence = urban_columbidae_distributions$presence_desc
  ),
 rbind(missing_eucdov_cities, rbind(missing_rocpig_cities, missing_spodov_cities))
)
```

```{r}
BIRDLIFE_URBAN_DISTRIBUTION = filename(BIRDLIFE_WORKING_OUTPUT_DIR, 'urban_distributions.csv')
write_csv(urban_distribution, BIRDLIFE_URBAN_DISTRIBUTION)
```

## Examine result
```{r}
urban_distribution = read_csv(BIRDLIFE_URBAN_DISTRIBUTION)
```

Species in Manchester's regional pool
```{r}
urban_distribution %>% filter(city_name == 'Manchester') 
```

Species in Miami's regional pool
```{r}
urban_distribution %>% filter(city_name == 'Miami') 
```

Cities with Common Woodpigeon in regional pool
```{r}
urban_distribution %>% filter(species_name == 'Columba palumbus')
```

Cities with Collared Dove in regional pool
```{r}
urban_distribution %>% filter(species_name == 'Streptopelia decaocto')
```

Regional pool sizes
```{r}
urban_distribution %>% group_by(city_id, city_name) %>% summarise(pool_size = n()) %>% arrange(pool_size)
```

