---
title: "Map Birdlife and eBird taxonomy to Jetz"
output: html_notebook
bibliography: ../ref.bib 
---

```{r}
source('../env.R')
```

We map to Jetz as our working taxonomy to allow us to use a clade tree.

# Map Birdlife and eBird

## Compare missing species
Read in Birdlife
```{r}
birdlife_tax = read_csv(filename(BIRDLIFE_WORKING_OUTPUT_DIR, 'taxonomy.csv'))
names(birdlife_tax) = c('birdlife_id', 'species_name', 'birdlife_common_name', 'alternative_name')
head(birdlife_tax)
```

Read in eBird
```{r}
ebird_tax = read_csv(filename(EBIRD_WORKING_OUTPUT_DIR, 'taxonomy.csv'))
names(ebird_tax) = c('ebird_id', 'ebird_category', 'ebird_species_name', 'ebird_common_name')
head(ebird_tax)
```
First create the root species name for all ebird records
```{r}
ebird_tax$species_name = word(ebird_tax$ebird_species_name, 1, 2, sep = ' ')
head(ebird_tax)
```

Species in eBird but not Birdlife
```{r}
ebird_tax[!(ebird_tax$species_name %in% birdlife_tax$species_name),]
```

Species in Birdlife but not eBird
```{r}
birdlife_tax[!(birdlife_tax$species_name %in% ebird_tax$species_name),]
```

Trying using Birdlife alternative name for renaming species
```{r}
birdlife_species_name_not_in_ebird = birdlife_tax[!(birdlife_tax$species_name %in% ebird_tax$species_name),]
birdlife_species_name_not_in_ebird[!(birdlife_species_name_not_in_ebird$alternative_name %in% ebird_tax$species_name),]
```

## Build initial taxonomy mapping with straight forward species
```{r}
ebird_birdlife_taxonomy_mapping = inner_join(
  birdlife_tax[, c('species_name', 'birdlife_id', 'birdlife_common_name')], 
  ebird_tax[, c('species_name', 'ebird_id', 'ebird_common_name', 'ebird_species_name')]
)
head(ebird_birdlife_taxonomy_mapping)
```

```{r}
mapped_by_alternative_name = inner_join(
  birdlife_species_name_not_in_ebird[, c('species_name', 'birdlife_id', 'birdlife_common_name', 'alternative_name')], 
  ebird_tax[, c('species_name', 'ebird_id', 'ebird_common_name', 'ebird_species_name')],
  by = c('alternative_name' = 'species_name')
)
mapped_by_alternative_name
```

```{r}
ebird_birdlife_taxonomy_mapping = rbind(
  ebird_birdlife_taxonomy_mapping,
  mapped_by_alternative_name[,c('species_name', 'birdlife_id', 'birdlife_common_name', 'ebird_id', 'ebird_common_name', 'ebird_species_name')]
)
```


## Species left to map
Species missing in Birdlife
```{r}
birdlife_tax[!(birdlife_tax$species_name %in% ebird_birdlife_taxonomy_mapping$species_name),]
```

Species missing in eBird
```{r}
ebird_tax[!(ebird_tax$ebird_species_name %in% ebird_birdlife_taxonomy_mapping$ebird_species_name),]
```

### Birdlife
Extinct Birdlife species

* Nesoenas duboisi

Absent in eBird

* Leptotila battyi: endemic to Panama, no records on eBird of this species or any Leptotila species in Panama.

Spilopelia suratensis

* https://avibase.bsc-eoc.org/species.jsp?lang=EN&avibaseid=AEF9E7D341798053
* https://ebird.org/species/spodov
* http://datazone.birdlife.org/species/factsheet/western-spotted-dove-spilopelia-suratensis

```{r}
birdlife_spilopelia_suratensis = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Spilopelia chinensis',c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Spilopelia suratensis', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_spilopelia_suratensis)
birdlife_spilopelia_suratensis
```

Patagioenas albipennis

* https://avibase.bsc-eoc.org/species.jsp?lang=EN&avibaseid=9DAE75390A64BCF5
* http://datazone.birdlife.org/species/factsheet/white-winged-pigeon-patagioenas-albipennis
* https://ebird.org/species/spwpig3/

```{r}
birdlife_patagioenas_albipennis = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Spilopelia chinensis',c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Patagioenas albipennis', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_patagioenas_albipennis)
birdlife_patagioenas_albipennis
```

Patagioenas albilinea

* https://avibase.bsc-eoc.org/species.jsp?lang=EN&avibaseid=975A928B4D624F19
* http://datazone.birdlife.org/species/factsheet/southern-band-tailed-pigeon-patagioenas-albilinea
* https://ebird.org/species/batpig1/
```{r}
birdlife_patagioenas_albilinea = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Patagioenas fasciata',c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Patagioenas albilinea', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_patagioenas_albilinea)
birdlife_patagioenas_albilinea
```
Alopecoenas hoedtii

* https://avibase.bsc-eoc.org/species.jsp?lang=EN&avibaseid=CD21DBC9B8F0385B
* https://ebird.org/species/wegdov1/

```{r}
birdlife_alopecoenas_hoedtii = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Pampusana hoedtii',c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Alopecoenas hoedtii', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_alopecoenas_hoedtii)
birdlife_alopecoenas_hoedtii
```
Alopecoenas jobiensis

* https://avibase.bsc-eoc.org/species.jsp?lang=EN&avibaseid=CD21DBC9B8F0385B
* https://ebird.org/species/wegdov1/

```{r}
birdlife_alopecoenas_hoedtii = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Pampusana hoedtii',c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Alopecoenas hoedtii', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_alopecoenas_hoedtii)
birdlife_alopecoenas_hoedtii
```

Alopecoenas jobiensis
```{r}
birdlife_alopecoenas_jobiensis = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Pampusana jobiensis',c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Alopecoenas jobiensis', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_alopecoenas_jobiensis)
birdlife_alopecoenas_jobiensis
```
Alopecoenas kubaryi
```{r}
birdlife_alopecoenas_kubaryi = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Pampusana kubaryi',c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Alopecoenas kubaryi', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_alopecoenas_kubaryi)
birdlife_alopecoenas_kubaryi
```
Alopecoenas xanthonurus
```{r}
birdlife_alopecoenas_xanthonurus = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Pampusana xanthonura',c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Alopecoenas xanthonurus', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_alopecoenas_xanthonurus)
birdlife_alopecoenas_xanthonurus
```
Alopecoenas erythropterus
```{r}
birdlife_alopecoenas_erythropterus = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Pampusana erythroptera',c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Alopecoenas erythropterus', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_alopecoenas_erythropterus)
birdlife_alopecoenas_erythropterus
```
Alopecoenas rubescens
```{r}
birdlife_alopecoenas_rubescens = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Pampusana rubescens',c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Alopecoenas rubescens', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_alopecoenas_rubescens)
birdlife_alopecoenas_rubescens
```
Alopecoenas beccarii	
```{r}
birdlife_alopecoenas_beccarii = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Pampusana beccarii',c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Alopecoenas beccarii', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_alopecoenas_beccarii)
birdlife_alopecoenas_beccarii
```
Alopecoenas johannae	
```{r}
birdlife_alopecoenas_johannae = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Pampusana beccarii [johannae Group]', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Alopecoenas johannae', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_alopecoenas_johannae)
birdlife_alopecoenas_johannae
```
Alopecoenas canifrons
```{r}
birdlife_alopecoenas_canifrons = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Pampusana canifrons', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Alopecoenas canifrons', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_alopecoenas_canifrons)
birdlife_alopecoenas_canifrons
```

Alopecoenas sanctaecrucis
```{r}
birdlife_alopecoenas_sanctaecrucis = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Pampusana sanctaecrucis', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Alopecoenas sanctaecrucis', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_alopecoenas_sanctaecrucis)
birdlife_alopecoenas_sanctaecrucis
```

Alopecoenas stairi
```{r}
birdlife_alopecoenas_stairi = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Pampusana stairi', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Alopecoenas stairi', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_alopecoenas_stairi)
birdlife_alopecoenas_stairi
```
Alopecoenas ferrugineus
```{r}
birdlife_alopecoenas_ferrugineus = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Pampusana ferruginea', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Alopecoenas ferrugineus', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_alopecoenas_ferrugineus)
birdlife_alopecoenas_ferrugineus
```
Alopecoenas salamonis
```{r}
birdlife_alopecoenas_salamonis = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Pampusana salamonis', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Alopecoenas salamonis', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_alopecoenas_salamonis)
birdlife_alopecoenas_salamonis
```
Otidiphaps aruensis

* https://avibase.bsc-eoc.org/species.jsp?lang=EN&avibaseid=F197D25F1AB50C3B
* https://ebird.org/species/phepig1/

```{r}
birdlife_otidiphaps_aruensis = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Otidiphaps nobilis aruensis', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Otidiphaps aruensis', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_otidiphaps_aruensis)
birdlife_otidiphaps_aruensis
```

Otidiphaps cervicalis
```{r}
birdlife_otidiphaps_cervicalis = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Otidiphaps nobilis cervicalis', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Otidiphaps cervicalis', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_otidiphaps_cervicalis)
birdlife_otidiphaps_cervicalis
```

Otidiphaps insularis
```{r}
birdlife_otidiphaps_insularis = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Otidiphaps nobilis insularis', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Otidiphaps insularis', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_otidiphaps_insularis)
birdlife_otidiphaps_insularis
```

Phapitreron nigrorum
```{r}
birdlife_phapitreron_nigrorum = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Phapitreron leucotis nigrorum', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Phapitreron nigrorum', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_phapitreron_nigrorum)
birdlife_phapitreron_nigrorum
```
Phapitreron brevirostris
```{r}
birdlife_phapitreron_brevirostris = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Phapitreron leucotis brevirostris/occipitalis', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Phapitreron brevirostris', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_phapitreron_brevirostris)
birdlife_phapitreron_brevirostris
```
Phapitreron maculipectus
```{r}
birdlife_phapitreron_maculipectus = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Phapitreron amethystinus maculipectus', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Phapitreron maculipectus', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_phapitreron_maculipectus)
birdlife_phapitreron_maculipectus
```
Phapitreron frontalis
```{r}
birdlife_phapitreron_frontalis = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Phapitreron amethystinus frontalis', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Phapitreron frontalis', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_phapitreron_frontalis)
birdlife_phapitreron_frontalis
```
Treron delalandii

* https://avibase.bsc-eoc.org/search.jsp?qstr=Treron+permagnus

```{r}
birdlife_treron_delalandii = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Treron calvus delalandii/granti', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Treron delalandii', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_treron_delalandii)
birdlife_treron_delalandii
```
Treron permagnus

* https://avibase.bsc-eoc.org/species.jsp?lang=EN&avibaseid=1EE907AC088D0B85

```{r}
birdlife_treron_permagnus = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Treron formosae permagnus/medioximus', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Treron permagnus', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_treron_permagnus)
birdlife_treron_permagnus
```
Ducula obiensis

* https://avibase.bsc-eoc.org/species.jsp?lang=EN&avibaseid=AAC237408FEB8125

```{r}
birdlife_ducula_obiensis = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Ducula basilica obiensis', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Ducula obiensis', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_ducula_obiensis)
birdlife_ducula_obiensis
```
Ducula salvadorii
```{r}
birdlife_ducula_salvadorii = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Ducula pinon salvadorii', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Ducula salvadorii', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_ducula_salvadorii)
birdlife_ducula_salvadorii
```
Ramphiculus epius
```{r}
birdlife_ramphiculus_epius = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Ptilinopus epius', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Ramphiculus epius', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_ramphiculus_epius)
birdlife_ramphiculus_epius
```
Ramphiculus subgularis	
```{r}
birdlife_ramphiculus_subgularis = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Ptilinopus subgularis', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Ramphiculus subgularis', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_ramphiculus_subgularis)
birdlife_ramphiculus_subgularis
```
Ramphiculus mangoliensis
```{r}
birdlife_ramphiculus_mangoliensis = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Ptilinopus mangoliensis', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Ramphiculus mangoliensis', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_ramphiculus_mangoliensis)
birdlife_ramphiculus_mangoliensis
```
Ramphiculus fischeri
```{r}
birdlife_ramphiculus_fischeri = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Ptilinopus fischeri', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Ramphiculus fischeri', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_ramphiculus_fischeri)
birdlife_ramphiculus_fischeri
```
Ramphiculus meridionalis

* https://avibase.bsc-eoc.org/species.jsp?lang=EN&avibaseid=E41C677F7619363F

```{r}
birdlife_ramphiculus_meridionalis = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Ptilinopus fischeri meridionalis', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Ramphiculus meridionalis', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_ramphiculus_meridionalis)
birdlife_ramphiculus_meridionalis
```
Ptilinopus temminckii

* https://avibase.bsc-eoc.org/species.jsp?lang=EN&avibaseid=722E8494A2E5BD3C

```{r}
birdlife_ptilinopus_temminckii = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Ptilinopus superbus temminckii', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Ptilinopus temminckii', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_ptilinopus_temminckii)
birdlife_ptilinopus_temminckii
```
Ptilinopus gestroi

* https://avibase.bsc-eoc.org/species.jsp?lang=EN&avibaseid=E0D24D45C31E80AB

```{r}
birdlife_ptilinopus_gestroi = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Ptilinopus ornatus gestroi', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Ptilinopus gestroi', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_ptilinopus_gestroi)
birdlife_ptilinopus_gestroi
```

Ptilinopus fasciatus

* https://avibase.bsc-eoc.org/species.jsp?lang=EN&avibaseid=9F3042D2364F20B7

```{r}
birdlife_ptilinopus_fasciatus = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Ptilinopus porphyraceus fasciatus', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Ptilinopus fasciatus', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, birdlife_ptilinopus_fasciatus)
birdlife_ptilinopus_fasciatus
```

### eBird
Extinct eBird species

* Pampusana norfolkensis

What hasn't been mapped?
```{r}
ebird_tax[!(ebird_tax$ebird_species_name %in% ebird_birdlife_taxonomy_mapping$ebird_species_name),]
```
Macropygia doreya

* https://avibase.bsc-eoc.org/species.jsp?avibaseid=DF44EED5
* https://avibase.bsc-eoc.org/species.jsp?lang=EN&avibaseid=FD5160925CD583C1

```{r}
ebird_macropygia_doreya = cbind(
  ebird_tax[ebird_tax$species_name == 'Macropygia doreya', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Macropygia amboinensis', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, ebird_macropygia_doreya)
ebird_macropygia_doreya
```

Macropygia cinnamomea

* https://avibase.bsc-eoc.org/species.jsp?lang=EN&avibaseid=F30B6A4781D2FA10

```{r}
ebird_macropygia_cinnamomea = cbind(
  ebird_tax[ebird_tax$species_name == 'Macropygia cinnamomea', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Macropygia emiliana', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, ebird_macropygia_cinnamomea)
ebird_macropygia_cinnamomea
```
Macropygia modiglianii

* https://avibase.bsc-eoc.org/species.jsp?lang=EN&avibaseid=3A905A46&sec=summary&ssver=1

```{r}
ebird_macropygia_modiglianii = cbind(
  ebird_tax[ebird_tax$species_name == 'Macropygia modiglianii', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Macropygia emiliana', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, ebird_macropygia_modiglianii)
ebird_macropygia_modiglianii
```
Pampusana beccarii beccarii
```{r}
ebird_pampusana_beccarii_beccarii = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Pampusana beccarii beccarii',c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Alopecoenas beccarii', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, ebird_pampusana_beccarii_beccarii)
ebird_pampusana_beccarii_beccarii
```


Ptilinopus fischeri fischeri/centralis
```{r}
ebird_ptilinopus_fischeri = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Ptilinopus fischeri fischeri/centralis', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Ramphiculus fischeri', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, ebird_ptilinopus_fischeri)
ebird_ptilinopus_fischeri
```

Ducula nicobarica

* https://avibase.bsc-eoc.org/species.jsp?lang=EN&avibaseid=2EB3A41049FE360A

```{r}
ebird_ducula_nicobarica = cbind(
  ebird_tax[ebird_tax$ebird_species_name == 'Ducula nicobarica', c('ebird_id', 'ebird_species_name', 'ebird_common_name')],
  birdlife_tax[birdlife_tax$species_name == 'Ducula aenea', c('birdlife_id', 'species_name', 'birdlife_common_name')]
)
ebird_birdlife_taxonomy_mapping = rbind(ebird_birdlife_taxonomy_mapping, ebird_ducula_nicobarica)
ebird_ducula_nicobarica
```
## Confirm all non-extinct species mapped
Species missing in Birdlife
```{r}
birdlife_tax[!(birdlife_tax$species_name %in% ebird_birdlife_taxonomy_mapping$species_name),]
```

Species missing in eBird
```{r}
ebird_tax[!(ebird_tax$ebird_species_name %in% ebird_birdlife_taxonomy_mapping$ebird_species_name),]
```

# Phylogentic tree using Open Tree of Life
